---
title: "imerim analysis"
author: "Nick R. Bachelder"
date: "9/10/2021"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(haven)
library(xlsx)
library(purrr)
library(readxl)
library(here)
setwd(here())
```


## pull data

```{r}
filenames <- list.files(paste(here(), "/aft25"), pattern="*.sas7bdat", full.names=TRUE)
ldf <- lapply(filenames, read_sas)


ldf
```

## merge data. This is messy but can be subsetted well.

```{r}
final_df <- data.frame(ldf[[1]])
i = 1
for (index in 2:length(ldf)) {
  merge_df <- data.frame(ldf[[index]])
  merge_columns <- intersect(colnames(final_df), colnames(merge_df))
  if (length(merge_columns) == 0) {
    next
  }
  final_df <- merge(final_df, merge_df, all = T, by = merge_columns)
  print(paste('Dataframe', i, 'done out of', length(ldf)))
  i = i + 1
}
```

## this is a list of relevant useful pages to subset master df

```{r}
unique(final_df$DataPageName)
```

Here find Race, Ethnicity, 

# get demographics
```{r}
registration <- final_df %>% filter(DataPageName == "Subject Enrollment" )
registration <- registration[colSums(!is.na(registration)) > 0] %>% select(subjectId, DATE_ON) %>% rename(enrolled_on = DATE_ON)

registration
```

# get regestration
```{r}
demo <- final_df %>% filter(DataPageName == 'Demographics')
demo <- demo[colSums(!is.na(demo)) > 0]
demo <- demo %>% select(subjectId, RACE, ETHNICITY)

demo
```




# get on-study
```{r}
### HPTUSEYN is history of hormone replacement thearapy, DCISLAT is laterality, 

onstudy <- final_df %>% filter(DataPageName == "On-Study")
onstudy <- onstudy[colSums(!is.na(onstudy)) > 0]
onstudy <- onstudy %>% select(subjectId, DCISLAT, HPTUSEYN, HPTLASTDT, HTUSECUR, MENOSTATRAND, HEIGHT, WEIGHT, ECOGPSBSL) %>% rename(lat = DCISLAT, hist_hormone_repl = HPTUSEYN, 
                                                                                                               dt_last_hormone_repl = HPTLASTDT, meno = MENOSTATRAND
                                                                                                               , ECOG = ECOGPSBSL, curr_hormone_rep = HTUSECUR)

onstudy
```



# get on-study left breast
```{r}
### ER is er status (positive if over 10%.)

onstudy_left <- final_df %>% filter(DataPageName == "On-Study: DCIS Disease Status (Left Breast)")
onstudy_left <- onstudy_left[colSums(!is.na(onstudy_left)) > 0]
onstudy_left <- onstudy_left %>% select(subjectId, ER, PGR) %>% rename(ER_left = ER, PGR_left = PGR)

onstudy_left
```


# get on-study right breast
```{r}
### HPTUSEYN is history of hormone replacement thearapy, DCISLAT is laterality, 

onstudy_right <- final_df %>% filter(DataPageName == "On-Study: DCIS Disease Status (Right Breast)")
onstudy_right <- onstudy_right[colSums(!is.na(onstudy_right)) > 0]
onstudy_right <- onstudy_right %>% select(subjectId, ER, PGR) %>% rename(ER_right = ER, PGR_right = PGR)

onstudy_right
```



# get prior surgery
```{r}
priorsurg <- final_df %>% filter(DataPageName == 'Summary of Prior DCIS Surgeries')
priorsurg <- priorsurg[colSums(!is.na(priorsurg)) > 0]
priorsurg <- priorsurg %>% mutate(prior_surg = ifelse(is.na(SURG_DT), 0, 1)) %>% select(subjectId, prior_surg, SURG_DT) %>% rename(prior_surg_dt = SURG_DT)

priorsurg
```



# get surgery
```{r}
surg <- final_df %>% filter(DataPageName == 'Surgical Summary') %>% select(subjectId, SURG_DT, SURGTYP, SURGLAT,DCISDZEXTENT,DCISWIDTH, ER, PGR,
                                                                           IHCHER2STAT, SLNDYN, SLNDTOTALN, SLNDPOSN, ALNDYN, ALNDAXILN, ALNDPOSN,
                                                                           NUMADDSURGREQ)
surg <- surg[colSums(!is.na(surg)) > 0]
surg1 <- surg %>% group_by(subjectId) %>% filter(SURG_DT == min(SURG_DT)) ## get each  patients first surgery
surg_count <- surg %>% group_by(subjectId) %>% count() %>% rename(surjery_count = n) ## count patient surgeries

surg_first <- merge(surg1, surg_count, by = 'subjectId', all = T) %>% mutate(multiple_surg = surjery_count > 1) %>% rename(surg_dt = SURG_DT,
                                                                                                                           surg_type = SURGTYP,
                                                                                                                           surg_lat = SURGLAT,
                                                                                                                           dcis_extent = DCISDZEXTENT,
                                                                                                                           dcis_margin = DCISWIDTH,
                                                                                                                           surg_er_status = ER,
                                                                                                                           surg_pgr_status = PGR,
                                                                                                                           surg_her2_status =
                                                                                                                             IHCHER2STAT,
                                                                                                                           sent_node_preform = SLNDYN,
                                                                                                                           sent_node_disect =
                                                                                                                             SLNDTOTALN,
                                                                                                                           pos_sent_node = SLNDPOSN,
                                                                                                                           axi_lymph_preform = 
                                                                                                                             ALNDYN,
                                                                                                                           axi_lymph_disect = 
                                                                                                                             ALNDAXILN,
                                                                                                                           pos_axi_lymph = 
                                                                                                                             ALNDPOSN,
                                                                                                                           additional_surg =
                                                                                                                             NUMADDSURGREQ)
surg_first
```



# get randomizations
```{r}
randomize <- final_df %>% filter(DataPageName == 'Randomization')
randomize <- randomize[colSums(!is.na(randomize)) > 0]
randomize <- randomize %>% select(subjectId, REGIME_NAME_DRV, DCISAGE, MICROCALC, GRADE) %>% rename(DCIS_margin = MICROCALC, DCIS_age = DCISAGE, DCIS_grade = GRADE, randomization_orig = REGIME_NAME_DRV)

randomize
```



# get assignment switches
```{r}
### SURGOPT means switched from AM to Surg and DECLNALLOCARM means switched from surgery to AM

switch <- final_df %>% filter(DataPageName == 'Change In Treatment/Management Approach (Including Allocation Refusal)')
switch <- switch[colSums(!is.na(switch)) > 0]
switch <- switch %>% mutate(switch_type = dplyr::case_when(SURGOPT == 'Yes' & DECLNALLOCARM == 'Yes' ~ 'Both',
                                          SURGOPT == 'Yes' & DECLNALLOCARM == '' ~ 'AM-->Surgery',
                                          SURGOPT == '' & DECLNALLOCARM == 'Yes' ~ 'Surgery-->AM',
                                          TRUE ~ 'NA')) %>%
  mutate(switch_type = na_if(switch_type, "NA"), switch = ifelse(is.na(switch_type), 0, 1),
         switch_date = ifelse(switch_type == 'AM-->Surgery', DCLNAMDT, DCLNTXRECMDT)) %>% select(subjectId, SURGOPT, DECLNALLOCARM, DCISCHNGE,  DCISNEW, switch_type, switch_date) %>%
  mutate(switch_date = as.POSIXct(switch_date, origin = "1970-01-01")) %>% rename(denied_AM_to_surg = SURGOPT,
                                                                                  denied_surg_to_AM = DECLNALLOCARM,
                                                                                  change_in_DCIS_reason_for_change = DCISCHNGE,
                                                                                  new_DCIS_reason_for_change = DCISNEW) %>% unique()

switch %>% filter(subjectId == 1356867)
```



Get needle biopsy
```{r}
biopsy <- final_df %>% filter(DataPageName == "Needle Biopsy")
biopsy <- biopsy[colSums(!is.na(biopsy)) > 0]
biopsy <- biopsy %>% select(subjectId, BIOPSYNDTXX, BREASTLATXX, ER, PGR, HER2STAT) %>% rename(biopsy_date = BIOPSYNDTXX, biopsy_lat = BREASTLATXX,
                                                                                    biopsy_er_status = ER, biopsy_pgr_status = PGR,
                                                                                    biopsy_her2_status = HER2STAT) %>% unique() 
biopsy 

```


Get thearapy treatment

```{r}
endocrine <- final_df %>% filter(DataPageName == 'Endocrine Therapy Treatment')
endocrine <- endocrine[colSums(!is.na(endocrine)) > 0] %>% select(subjectId, FRSTDOSEDURTX, LASTDOSEDURTX)
endocrine_yn <- unique(endocrine$subjectId)

endocrine ## dates are very inconsistant (discuss w terry)
```


get "Consent Withdrawal:  All Follow-Up"

```{r}
withdrawl <- final_df %>% filter(DataPageName == "End of Study (Study Withdrawal)")
withdrawl <- withdrawl[colSums(!is.na(withdrawl)) > 0]

withdrawl <- withdrawl %>% select(subjectId, ENDATDTHONSTUDY, ENDATCOM, ENDATFLLWCMPLT) %>% rename(withdraw_at_death = ENDATDTHONSTUDY, withdraw_w_FUP_comp = ENDATFLLWCMPLT,withdraw_reason_full = ENDATCOM)

withdrawl
```


try to find invasive status updates

"Patient Status"  = NO 
"Adverse Events: Other" = NO
"Deviations"= NO
"Recurrence, New Primary Breast Cancer, or DCIS"  
"Supporting Documentation: Recurrence, New Primary Breast Cancer, or DCIS"

IS THIS THE RIGHT DATE? SEEMS ODD THERES ONLY 46

```{r}
## invasive from recurrence folder

invasive_am <- final_df %>% filter(DataPageName == "Recurrence, New Primary Breast Cancer, or DCIS" ) 
invasive_am <- invasive_am[colSums(!is.na(invasive_am)) > 0] %>% select(subjectId, NEWMALIGDIAGDTXX) %>% rename(invasive_date = NEWMALIGDIAGDTXX) %>% mutate(from = 'am')

invasive_am

## invasive for surgery

invasive_surg <- final_df %>% filter(DataPageName == 'Surgical Summary') %>% select(subjectId, SURG_DT, INVASIVETUMOR)
invasive_surg <- invasive_surg[colSums(!is.na(invasive_surg)) > 0] %>% filter(INVASIVETUMOR == 'Yes') %>% rename(invasive_date = SURG_DT) %>% select(-INVASIVETUMOR) %>% mutate(from = 'surg')


invasive <- rbind(invasive_surg, invasive_am) %>% group_by(subjectId) %>% mutate(invasive_date = first(invasive_date, 
                                                                                           order_by = invasive_date)) %>% select(-from) %>% unique() ## choose first invasive date

```


### MERGE ALL THE VARIABLES TO MAKE COMPREHENSIVE DF

```{r}
merge_variables <- list(demo, registration, onstudy, onstudy_left, onstudy_right, priorsurg, surg_first, randomize, switch, biopsy, endocrine, withdrawl, invasive)
merge_fun <- function(x, y) {merge(x, y, by ='subjectId', all = T)}

full_df <- merge_variables %>% reduce(merge_fun) 
```

## THEN CLEAN EVERYTHING, CALCULATE TIME VALUES NECCECARY


## collapse into one row per patient with lists of variables

```{r}
library(lubridate)

sting_unique <- function(x) {toString(na.omit(unique(x)))}

full_df <- full_df %>%
    group_by(subjectId) %>% 
    summarise_each(funs(sting_unique)) 

full_df <- full_df %>% mutate(invasive_date = as.Date(invasive_date, format =  "%Y-%m-%d"), multiple_switch = ifelse(grepl( ',', switch_type, fixed = TRUE), 1, 0),
                              endorcine_therapy = ifelse(subjectId %in% endocrine_yn, 1, 0), enrolled_on = as.Date(enrolled_on, format =  "%Y-%m-%d"), 
                              time_to_invasive = invasive_date - enrolled_on) 

full_df
```



```{r}
write.csv(full_df, 'COMET_patient_summary.csv')
write.xlsx(full_df, 'COMET_patient_summary.xlsx')
```





```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```



```{r}

```



```{r}

```


```{r}

```

```{r}

```

```{r}

```




